Global Optimization of the Analogue Method by Means of Genetic Algorithms

### 2015-08-12 Optimization - Journal of Climate/template.tex
Discussion and conclusions.

\authors{Pascal Horton\correspondingauthor{Terranum, Rue de l'industrie 35 bis, 1030 Bussigny-pr\`{e}s-Lausanne, Switzerland.}}

@@

\affiliation{University of Lausanne, and Terranum,  Bussigny-pr\`{e}s-Lausanne, Switzerland}

@@

\email{pascal.horton@terranum.ch}

@@

\subsubsection{Results for the atmospheric circulation}

@@

The improvement of the CRPSS score relatively to the reference method on the atmospheric circulation is provided in Table \ref{table:scores} and is in average of 15.3 \% in calibration and 9.9 \% in validation. The score was also calculated for three precipitation thresholds: P\(\geq\)1 mm, P\(\geq\)0.1\(\cdot\)P10 and P\(\geq\)0.5\(\cdot\)P10, P10 being the daily precipitation with a 10~year return period (Table \ref{table:scores_thresholds_z4}). The score increases with the precipitation threshold: in average, respectively, 13.3\%, 15.4\% and 29.1\% in calibration and 7.9\%, 11.1\% and 34.5\% for the validation period. The optimized method improves thus even more days with significant precipitation than standard days.

@@

The scores show significant improvements for both the calibration and the validation period (Table \ref{table:scores}). Even more interestingly, the results for higher precipitation thresholds (Table \ref{table:scores_thresholds_z4}) show the largest improvements. This is of particular interest in the framework of flood forecasting.

@@

%Nous avons cherché à identifier quelle part du gain est attribuable à quel nouveau degré de liberté par rapport à la méthode M6. De manière générale, nous avons remarqué qu'il y a une forte dépendance entre ces paramètres (fenêtres spatiales, fenêtres temporelles et pondération des niveaux atmosphériques), et qu'il n'est pas recommandable de les considérer séparément. Par exemple, l'introduction des poids (non réoptimisés) dans les paramètres de la méthode M6 a un effet en moyenne nul, mais peut engendrer des pertes non négligeables sur les scores de certains groupements.

%An evaluation of all these parameters with the classic calibration requires multiple assessments of all possible combinations. In addition, these studies are often carried out with a spatial window of fixed size and a fixed number of analogues. So we hope to find solutions as good or better, but automatically.

@@

@@

When we introduce 2 predictors for the humidity analogy, the number of degrees of freedom raises to 42, and to 54 when we add 4 predictors.

%We tried first to optimize the humidity predictors while keeping the same parameters for the first level of analogy on the atmospheric circulation. The results were not really satisfying, mainly on the scores of the validation. -> POUR FLUX !

\subsubsection{Results for both analogy levels}
...

%Ces optimisations ont mené à des paramétrisations différentes de l'analogie de circulation par rapport à précédemment. Il existe donc bien des relations entre les paramètres des différents niveaux d'analogie, soit entre la circulation atmosphérique et le flux d'humidité. Ainsi, en connaissance de l'information du flux d'humidité, la méthode préférera une représentation différente de la circulation. Cet aspect n'a encore jamais été mis en évidence, car les outils habituels de calibration ne le permettent pas.

%Pour tenir compte des dépendances entre les paramètres des différents niveaux d'analogie, il est nécessaire d'utiliser une méthode d'optimisation globale telle que les AGs.

\subsubsection{Analysis}
...

@@

\begin{itemize}
	\item Using 4 atmospheric levels for the analogy of circulation. It seems to be the optimal number for the studied region. Beyond that value the validation score drops, revealing a lost in robustness due to over-parametrization.
	\item The automatic and joint optimization of all parameters: the analogues number, the choice of the atmospheric levels and the time windows, and the spatial windows. These parameters are highly interdependent, so we need to optimize them jointly in order to identify optimal combinations. Traditional calibration procedures based on a systematic assessment of every combination is no more possible when considering more than 2 atmospheric levels.
	\item The introduction of distinct spatial windows between atmospheric levels. Indeed, the synoptic circulation is characterized by features with very different scales depending on the height, and important information for predicting rainfall is not necessarily located in the same area from one level to another.
	\item The weighting of the analogy criteria between different atmospheric levels. It can be influenced by the variability of the geopotential with altitude, and the change of some levels significance with the targeted region. 
	\item The joint optimization of the circulation and humidity analogy levels, that are usually calibrated successively. We have been able to demonstrate that there is a dependency between the analogy levels, and that in order to approach the optimal parameters, we must consider them jointly.
\end{itemize}

@@

The convergence of parallel optimizations decreases when the method to optimize becomes more and more complex. The optimizer do not always converge to the exact global optimum, but to its surroundings. This is related to the fact that the optimization slows down when it gets closer to the global optimum, and that we have to stop it before the end due to the required processing time. The resulting parameters may sometimes present significant differences, even though the score is almost similar. Through some Monte-Carlo analyses of the parameter space properties of the analogue method, \citet{Horton2012a} showed that some parameters of the method have a wide range of acceptable values. The spatial windows, for example, can be larger than the optimal size without much impact on the score, while they cannot be smaller \citep[see also][]{Bontron2004}. We could also observe that the choice of the atmospheric level is not a parameter as discrete as we would have thought, and that choosing another level may have reduced consequences on the performance. This is particularly true for higher atmospheric levels and can be more critical for lower layers. It was thus interesting to sometimes get several sets of near-optimal parameters, but with some nuances, in order to have an idea of the sensitivity of the parameters for a given region, and to compare the score on the validation period. In this regard, a cross-validation approach may be recommended.

We tried to optimize the preselection period (the 4-months window) jointly with the other parameters, but we did not get improvement. We also tried optimizing the humidity flux, which is composed of the humidity index multiplied with the wind flux. However, the results were not as good as when we considered the humidity index alone. This may be related to the fact that the optimizer tries to provide the best analogy of the atmospheric circulation in the first place, which makes the wind information less relevant in the second level of analogy.

@@

% Cet aspect a été confirmé par l'optimisation, qui propose de petites fenêtres pour les niveaux supérieurs, et de grandes fenêtres pour les altitudes proches du sol. Notre hypothèse est que l'information des gradients des couches supérieures au-dessus du bassin suffit à résumer l'orientation et la vitesse des masses d'air d'intérêt, qui sont potentiellement liées à un forçage par-dessus la barrière orographique. De manière complémentaire, les fenêtres plus étendues pour les basses couches cherchent à intégrer la structure de la circulation, car celle-ci renseigne sur les centres actifs et sur le chemin qu'a parcouru la masse d'air éventuellement chargée en humidité avant d'arriver sur la zone d'intérêt.

@@

We could observe strong dependencies between the parameters of the analogue method. Thus, the classic calibration, which optimizes the parameters successively, may not lead to the optimal combination. Moreover, it contains several manual systematic assessment, such as the choice of the atmospheric levels and the temporal windows.

In order to automatically optimize the analogue method, we evaluated the genetic algorithms, a global optimization technique, that is able to solve complex problems. Given the large number of existing operators and options, we had to evaluate systematically multiple variants in order to identify which operators are important, and which variant works best for the analogue method. We could thus identify that the mutation operator is a key element for our application, and provided new variants that were found to be efficient. Recommendations were established for a relevant use of GAs for the optimization of the analogue method.

GAs provide parametrizations of the analogue method that exceed performance results of classic calibration. In addition, they are used to select the atmospheric levels and the time windows automatically, which was not possible with traditional tools and which can save a considerable amount of human time. A great advantage of a global optimization is its ability to approach or reach optimal parameter values when they are considered jointly.

@@

For our case study, there seems to be an optimum number of atmospheric levels to consider for the analogy of circulation, which is four, before losing performance in validation. We have also been able to improve the analogy of circulation by introducing a weighting between atmospheric levels, and considering independent spatial windows between levels.

However, there is a risk of over-parametrization with this type of approach. It is therefore important to always evaluate the proposed parameters over an independent period for validation.

%\item Mutation non uniforme: cet opérateur présente potentiellement le meilleur taux de convergence, pour autant qu'il soit bien paramétré. En revanche, si l'évolution de la population n'a pas été correctement estimée au préalable, et que nous choisissons une mauvaise valeur parmi ses trois paramètres, l'optimisation peut être médiocre. Son potentiel est donc prometteur, mais l'opérateur est difficile à maîtriser, car les paramètres sont nombreux et leur valeur dépend du problème à optimiser.
%\item Chromosome de rayons de recherche adaptatifs: nous avons élaboré le chromosome de rayons de recherche adaptatifs en nous inspirant du chromosome de taux de mutation adaptatifs. Notre version est plus performante, car elle intègre une notion de distance dans l'espace des paramètres, qui décroît avec l'évolution de la population, améliorant ainsi la convergence. Une grande force de cet opérateur est qu'il ne nécessite aucun paramètre, il s'autocontrôle et évolue avec la population, ce qui le rend très robuste.
%\item Mutation multi-échelles: la mutation multi-échelles est une nouvelle approche qui ne cherche pas une évolution de l'opérateur de mutation, mais qui effectue des recherches globales, régionales et locales en tout temps. Il est ainsi constamment en train d'explorer l'espace des paramètres complet et d'exploiter l'information locale, les deux aspects importants des algorithmes génétiques. Il ne nécessite qu'un seul paramètre, qui est le taux de mutation. Celui-ci doit cependant être bien choisi, et cette valeur change en fonction de la méthode à optimiser.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ACKNOWLEDGMENTS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\acknowledgments
Thanks to Hamid Hussain-Khan of the University of Lausanne for his help and availability, and for the intensive use of the cluster he is in charge of.

@@

\begin{table}[htbp]
	\footnotesize
	\caption{Improvement (\%) of the CRPSS of the three methods. For z4, the improvement is relative to the reference method on the atmospheric circulation. For z4-hi2 and z4-hi4, the improvement is relative to the reference method on the circulation and humidity variables.}
	\begin{center}
		\begin{tabular}{rrarara}
			\hline 
			\textbf{ID} & \multicolumn{2}{c}{\textbf{z4}} & \multicolumn{2}{c}{\textbf{z4-hi2}} & \multicolumn{2}{c}{\textbf{z4-hi4}} \\ 
			\hline 
			& calib & valid & calib & valid & calib & valid \\ 
			\hline 
			\textbf{1} & 13.6 & 13.0 & 11.9 & 9.9 & 12.8 & 9.6 \\ \hline
			\textbf{2} & 13.3 & 7.5 & 11.2 & 9.4 & 12.5 & 10.8 \\ \hline
			\textbf{3} & 12.2 & 9.0 & 13.2 & 12.0 & 14.1 & 13.5 \\ \hline
			\textbf{4} & 14.2 & 10.5 & 12.4 & 10.6 & 14.5 & 12.5 \\ \hline
			\textbf{5} & 14.5 & 8.8 & 12.5 & 9.1 & 14.1 & 10.0 \\ \hline
			\textbf{6} & 13.3 & 8.6 & 11.5 & 8.2 & 14.1 & 8.4 \\ \hline
			\textbf{7} & 22.8 & 10.6 & 19.7 & 15.7 & 21.7 & 16.3 \\ \hline
			\textbf{8} & 20.7 & 11.9 & 19.2 & 13.5 & 20.8 & 14.0 \\ \hline
			\textbf{9} & 14.4 & 9.5 & 15.8 & 15.2 & 16.7 & 14.1 \\ \hline
			\textbf{10} & 14.2 & 9.4 & 12.9 & 11.2 & 14.6 & 11.8 \\ \hline
			\textbf{av.} & \textbf{15.3} & \textbf{9.9} &\textbf{14.0} & \textbf{11.5} & \textbf{15.6} & \textbf{12.1} \\ \hline 
		\end{tabular} 
	\end{center}
	\label{table:scores}
\end{table}

@@

\begin{table}[htbp]
	\footnotesize
	\caption{Parameters of the z4-hi2 method (atmospheric circulation on 4 levels and humidity index on 2 levels), optimized for the Chablais region. The columns are the followings: L = level, V = variable, H = hour, D = domain, W = weight, C = criteria, N = number of analogues.}
	\begin{center}
		\begin{tabular}{ccccccc}
			\hline \textbf{L} & \textbf{V} & \textbf{H} & \textbf{D} & \textbf{W} & \textbf{C} & \textbf{N} \\ 
			\hline 
			0 & \multicolumn{6}{l}{$\pm 60$ days around the target date} \\
			\hline 
			\multirow{8}{*}{1} &  \multirow{2}{*}{Z300} & \multirow{2}{*}{12~h} & 0.0 - 15.0 \degree E & \multirow{2}{*}{23\%} & \multirow{8}{*}{S1} & \multirow{8}{*}{74} \\
			& & & 42.5 - 55.0 \degree N & & & \\ 
			& \multirow{2}{*}{Z500} & \multirow{2}{*}{30~h} & 0.0 - 10.0 \degree E & \multirow{2}{*}{25\%} & & \\ 
			& & & 40.0 - 45.0 \degree N & & & \\ 
			& \multirow{2}{*}{Z850} & \multirow{2}{*}{12~h} & 0.0 - 10.0 \degree E & \multirow{2}{*}{26\%} & & \\ 
			& & & 42.5 - 47.5 \degree N & & & \\ 
			& \multirow{2}{*}{Z1000} & \multirow{2}{*}{30~h} & 0.0 - 20.0 \degree E & \multirow{2}{*}{26\%} & & \\ 
			& & & 40.0 - 47.5 \degree N & & & \\ 
			\hline 
			\multirow{4}{*}{2} & TPW & \multirow{2}{*}{12~h} & 5.0 - 10.0 \degree E & \multirow{2}{*}{57\%} & \multirow{4}{*}{RMSE} & \multirow{4}{*}{25} \\
			& *RH700 & & 45.0 - 47.5 \degree N & & & \\ 
			& TPW & \multirow{2}{*}{24~h} & 5.0 - 10.0 \degree E & \multirow{2}{*}{43\%} & & \\ 
			& *RH700 & & 45.0 - 47.5 \degree N & & & \\ 
			\hline 
		\end{tabular} 
	\end{center}
	\label{table:params_GA_z4_hi2}
\end{table}

>>>>

\authors{Pascal Horton\correspondingauthor{Terranum, Rue de l'industrie 35 bis, 1030 Bussigny, Switzerland.}}

@@

\affiliation{University of Lausanne and Terranum, Lausanne, Switzerland}

@@

\email{pascal.horton@alumnil.unil.ch}

@@

\subsubsection{Results for the z4 method}

@@

The resulting CRPSS scores are provided in Table \ref{table:scores} and is in average 35.8 \% in calibration and 35.5 \% in validation. The improvement of the CRPSS score relatively to the reference method on the atmospheric circulation is provided in Table \ref{table:scores_diff} and is in average 15.3 \% in calibration and 9.9 \% in validation. The score was also calculated for three precipitation thresholds: P\(\geq\)1 mm, P\(\geq\)0.1\(\cdot\)P10 and P\(\geq\)0.5\(\cdot\)P10, P10 being the daily precipitation with a 10~year return period (Table \ref{table:scores_thresholds_z4}). The score increases with the precipitation threshold: in average, respectively, 13.3\%, 15.4\% and 29.1\% in calibration and 7.9\%, 11.1\% and 34.5\% for the validation period. The optimized method improves thus even more days with significant precipitation than standard days.

@@

The scores show significant improvements for both the calibration and the validation period (Table \ref{table:scores_diff}). Even more interestingly, the results for higher precipitation thresholds (Table \ref{table:scores_thresholds_z4}) show the largest improvements. This is of particular interest in the framework of flood forecasting.

@@

@@

Two methods were assessed: one with 2 humidity predictors (humidity index on 2 atmospheric levels or at 2 different hours) named z4-hi2, and one with 4 humidity predictors named z4-hi4. When we introduce 2 predictors for the humidity analogy, the number of degrees of freedom raises to 42, and to 54 when we add 4 predictors.

@@

\subsubsection{Results for z4-hi2 and z4-hi4 methods}

As previously, the optimized parameters differ from one region to another, and this event to a greater extent. Detailed examples are still provided for the Swiss Chablais region in Table \ref{table:params_GA_z4_hi2} for z4-hi2 and in Table \ref{table:params_GA_z4_hi4} for z4-hi4. The resulting spatial windows are displayed in Figure \ref{fig:spatial_windows_z4-hi2} for z4-hi2, along with the selected atmospheric levels for the circulation analogy (Table \ref{table:levels_GA_z4_hi2}) and the humidity analogy (Table \ref{table:levels_GA_z4_hi2_H}). Correspondingly, for the method z4-hi4, the spatial windows are shown in Figure \ref{fig:spatial_windows_z4-hi4} and the selected atmospheric levels are presented in Tables \ref{table:levels_GA_z4_hi4} and \ref{table:levels_GA_z4_hi4_H}.

The CRPSS scores of the optimized methods are provided in Table \ref{table:scores} and amounts to slightly more than 40 \% for both methods and for both periods. This results in a relative improvements that ranges from 11.5 \% to 15.6 \% compared to the reference method on the humidity analogy (Table \ref{table:scores_diff}). As for the method z4, methods z4-hi2 and z4-hi4 present improvements of the forecast of significant rainfall (thresholds P\(\geq\)1 mm, P\(\geq\)0.1\(\cdot\)P10 and P\(\geq\)0.5\(\cdot\)P10). The improvement are relatively similar for z4-hi2 (Table \ref{table:scores_thresholds_z4-hi2}) and z4-hi4 (Table \ref{table:scores_thresholds_z4-hi4}), with slightly superior scores for the latter on small precipitation (P\(\geq\)1 mm) and extremes (P\(\geq\)0.5\(\cdot\)P10).

The parameters cross-compatibility has also been assessed for the methods with humidity variables, and are shown in Tables \ref{table:crossing_z4-hi2_calib} and \ref{table:crossing_z4-hi2_valid} for the method z4-hi2, and Tables \ref{table:crossing_z4-hi4_calib} and \ref{table:crossing_z4-hi4_valid} for z4-hi4.

When optimizing a method made of 2 levels of analogy, the introduction of humidity variables in the second level has an influence on the parameter values of the first level. This is first visible on the choice of the atmospheric levels for the circulation analogy, where the 1000~hPa that was previously systematically selected twice (Table \ref{table:levels_GA_z4}) is now less chosen (once or even not at all) for both za-hi2 (Table \ref{table:levels_GA_z4_hi2}) and za-hi4 (Table \ref{table:levels_GA_z4_hi4}). There is indeed a shift of the previously selected 1000~hPa for higher levels, which is even slightly stronger with 4 humidity predictors than with 2. This change is likely due to the fact that when considering only the circulation analogy, the method tries to take into account information that can serve as proxy for humidity assessment, whereas it doesn't need it with the humidity index. This aspect has never been demonstrated, as usual calibration tools do not allow it. It can only be assessed by a global optimization technique that can tune jointly both levels of analogy.

The selected levels for the analogy on the humidity index is strongly centred around the 700~hPa and 600~hPa. No other value has been selected when considering 2 atmospheric levels (Table \ref{table:levels_GA_z4_hi2_H}), and when considering 4 levels, the 850~hPa is quite often chosen, and the 500~hPa more rarely. However, even is this latter method, the 700~hPa and 600~hPa still hold 78 \% of the selected levels. It is thus more efficient to consider one of this level multiple times at different hours rather than using another atmospheric level. Besides, the optimizer never chose the same atmospheric level at the same hour for any variable.

The choice of the temporal windows for the atmospheric circulation is similar to the preceding optimization (in the order of increasing pressure: 12~h, 24/30~h, 12~h, 30~h), but sometimes with a bit more variability. When it comes to the humidity analogy, there is a clear trend to select 12~h and 24~h when considering 2 predictors. The z4-hi4 method presents more variability on the choice of the temporal windows for the humidity index.

The optimized spatial windows on the atmospheric circulation have also changed (Figures \ref{fig:spatial_windows_z4-hi2} and \ref{fig:spatial_windows_z4-hi4}). The very large domains on the upper level of the 4 first groupments are not present anymore, and more variability can be observed. The selected points for the analogy on the humidity index are always located nearby the catchment, including 1 of the nearest point from the reanalysis dataset, and the spatial windows are relatively small. Thus, for our case study, there is no need to look for distant humidity information and the search could be reduced to a smaller domain.

The CRPSS scores were improved by considering the humidity information (Table \ref{table:scores}). The optimized methods also perform better than the reference method based on the humidity index (Table \ref{table:scores_diff}). However, there is no drastic difference between both z4-hi2 and z4-hi4 methods, which suggests that considering 4 humidity predictors is not absolutely necessary. When it comes to improvements for days with precipitation above our 3 thresholds (P\(\geq\)1 mm, P\(\geq\)0.1\(\cdot\)P10 and P\(\geq\)0.5\(\cdot\)P10), the conclusion is the same as before, that is a significant improvement of the forecast, mainly for heavy rainfall.

The analysis of the parameters cross-compatibility (Tables \ref{table:crossing_z4-hi2_calib} to \ref{table:crossing_z4-hi4_valid}) is also very similar to the one on the circulation analogy only. The same pattern can be observed, with a drop of performance for the regions under different meteorological influences. However, the losses of performance are globally more important than before, suggesting that more complex methods with humidity variables are less transposable to another region, even though it is location within the same grid of the reanalysis dataset.

@@

\begin{itemize}
	\item Using 4 atmospheric levels for the analogy of circulation. It seems to be the optimal number for the studied region. Beyond that value the validation score drops, revealing a lost in robustness due to over-parametrization.
	\item The automatic and joint optimization of all parameters: the analogues number, the choice of the atmospheric levels and the time windows, and the spatial windows. These parameters are highly interdependent, so we need to optimize them jointly in order to identify optimal combinations. Traditional calibration procedures based on a systematic assessment of every combination is no more possible when considering more than 2 atmospheric levels.
	\item The introduction of distinct spatial windows between atmospheric levels. Indeed, the synoptic circulation is characterized by features with very different scales depending on the height, and important information for predicting rainfall is not necessarily located in the same area from one level to another. The optimized spatial windows are consistent in between the groupments.
	\item The weighting of the analogy criteria between different atmospheric levels. It can be influenced by the variability of the geopotential with altitude, and the change of some levels significance with the targeted region. There is a trend for the weighting of circulation predictors to decrease with the increase in pressure, as one can see in Figure \ref{fig:levels_weights} for the method z4, and in Figure \ref{fig:levels_weights_average} for the averages over the 3 methods. However, the values stay around equity. This may not be the most influencing factor, and we may suggest to remove it first when trying to reduce the number of degrees of freedom.
	\item The joint optimization of the circulation and humidity analogy levels, that are usually calibrated successively. We have been able to demonstrate that there is a dependency between the analogy levels, and that in order to approach the optimal parameters, we must consider them jointly.
\end{itemize}

@@

The convergence of parallel optimizations decreases when the method to optimize becomes more and more complex. The optimizer do not always converge to the exact global optimum, but to its surroundings. This is related to the fact that the optimization slows down when it gets closer to the global optimum, and that we have to stop it before the end, due to the required processing time (see Figure \ref{fig:evolution} as example). The resulting parameters may sometimes present significant differences, even though the score is almost similar. Through some Monte-Carlo analyses of the parameter space properties of the analogue method, \citet{Horton2012a} showed that some parameters of the method have a wide range of acceptable values. The spatial windows, for example, can be larger than the optimal size without much impact on the score, while they cannot be smaller \citep[see also][]{Bontron2004}. We could also observe that the choice of the atmospheric level is not a parameter as discrete as we would have thought, and that choosing another level may have reduced consequences on the performance. This is particularly true for higher atmospheric levels and can be more critical for lower layers. It was thus interesting to sometimes get several sets of near-optimal parameters, but with some nuances, in order to have an idea of the sensitivity of the parameters for a given region, and to compare the score on the validation period. In this regard, a cross-validation approach may be recommended.

We tried to optimize the preselection period (the 4-months window) jointly with the other parameters, but we did not get improvement. We also tried optimizing the humidity flux, which is composed of the humidity index multiplied with the wind flux. However, the results were not better than when we considered the humidity index alone. This may be related to the fact that the optimizer tries to provide the best analogy of the atmospheric circulation in the first place, which makes the wind information less relevant in the second level of analogy.

As we could observe, methods with a higher complexity that integrate humidity predictors are less transposable than simpler ones. We also noticed in another unpublished work, that it is by far better to optimize for 2 regions jointly rather than to optimize on one and to apply its parametrization to the other. Finally, the creation of the groupments is an important stage and should be handled with care. Indeed, as we could observe, the geographical proximity is not always the leading factor to define a groupment. For example, the Southeast ridges groupment do not behave like its surrounding and differ in its parametrization, due to different leading meteorological influences.

@@

@@

In order to automatically optimize the analogue method, we evaluated the genetic algorithms, a global optimization technique, that is able to solve complex problems. Given the large number of existing operators and options, we had to evaluate systematically multiple variants in order to identify which operators are important, and which variant works best for the analogue method. We could thus identify that the mutation operator is a key element for our application, and provided new variants that were found to be efficient, such as the Chromosome of adaptive search radius that seems to be very reliable (no control parameter) and efficient . Recommendations were established for a relevant use of GAs for the optimization of the analogue method.

The parameters resulting from the optimization are very consistent for the analogy of the atmospheric circulation, in terms of choice of the pressure levels, and the temporal and spatial windows. There are clear trends or even identical results for regions under similar meteorological influences, which confirm that the optimized parameters are coherent, despite an eventual first impression of a great variability in the spatial windows. When adding humidity variables, the results show a higher variability, but remains highly acceptable and coherent. Indeed, the choice of the pressure level for the humidity index is for example very constant.

We could observe strong dependencies between the parameters of the analogue method. Thus, the classic calibration, which optimizes the parameters successively, may not lead to the optimal combination. Moreover, it contains several manual systematic assessment, such as the choice of the atmospheric levels and the temporal windows. GAs, however, can select the atmospheric levels and the time windows automatically, which can save a considerable amount of human time. A great advantage of a global optimization is its ability to approach or reach optimal parameter values when they are considered jointly.

@@

For our case study, there seems to be an optimum number of atmospheric levels to consider for the analogy of circulation, which is four, before losing performance in validation. We have also been able to improve the analogy of circulation by introducing a weighting between atmospheric levels, and considering independent spatial windows between levels. However, there is a risk of over-parametrization with this type of approach. It is therefore important to always evaluate the proposed parameters over an independent period for validation.

GAs provide parametrizations of the analogue method that exceed performance results of classic calibration. In addition, we could observe that the forecasts for days with strong precipitation were significantly more improved, which is clearly interesting in the framework of flood forecasting.

Perspective .........

%\item Chromosome de rayons de recherche adaptatifs: nous avons élaboré le chromosome de rayons de recherche adaptatifs en nous inspirant du chromosome de taux de mutation adaptatifs. Notre version est plus performante, car elle intègre une notion de distance dans l'espace des paramètres, qui décroît avec l'évolution de la population, améliorant ainsi la convergence. Une grande force de cet opérateur est qu'il ne nécessite aucun paramètre, il s'autocontrôle et évolue avec la population, ce qui le rend très robuste.
%\item Mutation multi-échelles: la mutation multi-échelles est une nouvelle approche qui ne cherche pas une évolution de l'opérateur de mutation, mais qui effectue des recherches globales, régionales et locales en tout temps. Il est ainsi constamment en train d'explorer l'espace des paramètres complet et d'exploiter l'information locale, les deux aspects importants des algorithmes génétiques. Il ne nécessite qu'un seul paramètre, qui est le taux de mutation. Celui-ci doit cependant être bien choisi, et cette valeur change en fonction de la méthode à optimiser.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ACKNOWLEDGMENTS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\acknowledgments
Thanks to Hamid Hussain-Khan of the University of Lausanne for his help and availability, and for the intensive use of the cluster he is in charge of. Thanks to Renaud Marty for his fruitful collaboration over the years.

@@

\begin{table}[htbp]
	\footnotesize
	\caption{CRPSS score (\%) of the three optimized methods.}
	\begin{center}
		\begin{tabular}{rrarara}
			\hline 
			\textbf{ID} & \multicolumn{2}{c}{\textbf{z4}} & \multicolumn{2}{c}{\textbf{z4-hi2}} & \multicolumn{2}{c}{\textbf{z4-hi4}} \\ 
			\hline 
			& calib & valid & calib & valid & calib & valid \\ 
			\hline 
			\textbf{1} & 40.3 & 38.5 & 45.5 & 43.8 & 45.9 & 43.6 \\ \hline
			\textbf{2} & 37.8 & 34.9 & 42.4 & 40.8 & 42.9 & 41.3 \\ \hline
			\textbf{3} & 36.6 & 33.9 & 41.4 & 39.0 & 41.8 & 39.5 \\ \hline
			\textbf{4} & 31.2 & 30.8 & 36.1 & 36.4 & 36.8 & 37.1 \\ \hline
			\textbf{5} & 34.4 & 36.5 & 38.6 & 41.4 & 39.2 & 41.8 \\ \hline
			\textbf{6} & 37.0 & 39.9 & 41.0 & 44.5 & 41.9 & 44.6 \\ \hline
			\textbf{7} & 30.0 & 30.1 & 33.8 & 34.6 & 34.4 & 34.8 \\ \hline
			\textbf{8} & 37.7 & 40.1 & 40.7 & 43.1 & 41.2 & 43.3 \\ \hline
			\textbf{9} & 33.1 & 30.0 & 38.4 & 36.6 & 38.7 & 36.3 \\ \hline
			\textbf{10} & 40.0 & 40.0 & 42.6 & 42.9 & 43.3 & 43.2 \\ \hline
			\textbf{av.} & \textbf{35.8} & \textbf{35.5} &\textbf{40.1} & \textbf{40.3} & \textbf{40.6} & \textbf{40.5} \\ \hline 
		\end{tabular} 
	\end{center}
	\label{table:scores}
\end{table}

\begin{table}[htbp]
	\footnotesize
	\caption{Improvement (\%) of the CRPSS of the three methods. For z4, the improvement is relative to the reference method on the atmospheric circulation. For z4-hi2 and z4-hi4, the improvement is relative to the reference method on the circulation and humidity variables.}
	\begin{center}
		\begin{tabular}{rrarara}
			\hline 
			\textbf{ID} & \multicolumn{2}{c}{\textbf{z4}} & \multicolumn{2}{c}{\textbf{z4-hi2}} & \multicolumn{2}{c}{\textbf{z4-hi4}} \\ 
			\hline 
			& calib & valid & calib & valid & calib & valid \\ 
			\hline 
			\textbf{1} & 13.6 & 13.0 & 11.9 & 9.9 & 12.8 & 9.6 \\ \hline
			\textbf{2} & 13.3 & 7.5 & 11.2 & 9.4 & 12.5 & 10.8 \\ \hline
			\textbf{3} & 12.2 & 9.0 & 13.2 & 12.0 & 14.1 & 13.5 \\ \hline
			\textbf{4} & 14.2 & 10.5 & 12.4 & 10.6 & 14.5 & 12.5 \\ \hline
			\textbf{5} & 14.5 & 8.8 & 12.5 & 9.1 & 14.1 & 10.0 \\ \hline
			\textbf{6} & 13.3 & 8.6 & 11.5 & 8.2 & 14.1 & 8.4 \\ \hline
			\textbf{7} & 22.8 & 10.6 & 19.7 & 15.7 & 21.7 & 16.3 \\ \hline
			\textbf{8} & 20.7 & 11.9 & 19.2 & 13.5 & 20.8 & 14.0 \\ \hline
			\textbf{9} & 14.4 & 9.5 & 15.8 & 15.2 & 16.7 & 14.1 \\ \hline
			\textbf{10} & 14.2 & 9.4 & 12.9 & 11.2 & 14.6 & 11.8 \\ \hline
			\textbf{av.} & \textbf{15.3} & \textbf{9.9} &\textbf{14.0} & \textbf{11.5} & \textbf{15.6} & \textbf{12.1} \\ \hline 
		\end{tabular} 
	\end{center}
	\label{table:scores_diff}
\end{table}

@@

\begin{table}[htbp]
	\footnotesize
	\caption{Parameters of the z4-hi2 method (atmospheric circulation on 4 levels and humidity index on 2 levels), optimized for the Chablais region. Same conventions as Table \ref{table:params_GA_z4}.}
	\begin{center}
		\begin{tabular}{ccccccc}
			\hline \textbf{L} & \textbf{V} & \textbf{H} & \textbf{D} & \textbf{W} & \textbf{C} & \textbf{N} \\ 
			\hline 
			0 & \multicolumn{6}{l}{$\pm 60$ days around the target date} \\
			\hline 
			\multirow{8}{*}{1} &  \multirow{2}{*}{Z300} & \multirow{2}{*}{12~h} & 0.0 - 15.0 \degree E & \multirow{2}{*}{23\%} & \multirow{8}{*}{S1} & \multirow{8}{*}{74} \\
			& & & 42.5 - 55.0 \degree N & & & \\ 
			& \multirow{2}{*}{Z500} & \multirow{2}{*}{30~h} & 0.0 - 10.0 \degree E & \multirow{2}{*}{25\%} & & \\ 
			& & & 40.0 - 45.0 \degree N & & & \\ 
			& \multirow{2}{*}{Z850} & \multirow{2}{*}{12~h} & 0.0 - 10.0 \degree E & \multirow{2}{*}{26\%} & & \\ 
			& & & 42.5 - 47.5 \degree N & & & \\ 
			& \multirow{2}{*}{Z1000} & \multirow{2}{*}{30~h} & 0.0 - 20.0 \degree E & \multirow{2}{*}{26\%} & & \\ 
			& & & 40.0 - 47.5 \degree N & & & \\ 
			\hline 
			\multirow{4}{*}{2} & TPW & \multirow{2}{*}{12~h} & 5.0 - 10.0 \degree E & \multirow{2}{*}{57\%} & \multirow{4}{*}{RMSE} & \multirow{4}{*}{25} \\
			& *RH700 & & 45.0 - 47.5 \degree N & & & \\ 
			& TPW & \multirow{2}{*}{24~h} & 5.0 - 10.0 \degree E & \multirow{2}{*}{43\%} & & \\ 
			& *RH700 & & 45.0 - 47.5 \degree N & & & \\ 
			\hline 
		\end{tabular} 
	\end{center}
	\label{table:params_GA_z4_hi2}
\end{table}
